---
author: "Alireza Sheikh-Zadeh, PhD"
output:
  word_document: default
  html_document: default
title: "R Assignment 2"
---

Document format: Follow the instructions given on the web page. Always review your solution word document before submission.

Plagiarism: You are not allowed to share your write-up with your peers. It's okay to advise your peers about how to solve a problem, but you never share your own write-up.

Problem 1: 28 points

Problem 2: 52 points 

Format: 20 points

# Problem 1 (28 points)
Suppose you roll a die twice in succession, getting $X_1$ and $X_2$. Then divide them, getting $Y=\frac{X_1}{X_2}$. Thus, $Y$ is discrete, ranging from a minimum of 1/6 to a maximum of 6. In R, we can generate all possible values for $Y$ as follows.


```{r}
# run this chunk of code:
X1 = c(1,2,3,4,5,6) # possible outcome of the first try
X2 = c(1,2,3,4,5,6) # possible outcome of the second try

Y = c() # defining an empty vector Y
# Then we divide each element of X1 by all value of X2 and record them in Y vector
for (i in X1) {
  for (j in X2) {
     Y = c(Y, i/j) # inserting X1/X2 into Y vector
  }
}
# Now Y is ready to be used in part a
```


a. Report the sorted list of $Y$ values, then Report the mean of the $Y$. Present all values by 3 decimal places. (6 points)


Using the sort() function in R we sorted Y as seen below from lowest to greatest.  We then used the mean() function to calculate the mean or expected value of Y where E(Y) = mu = 1.429.


```{r}
# Hint: you can use sort() and mean() functions in R. 
round(sort(Y),3)
round(mean(Y),3)

```

b. Simulate 10000 (or more) i.i.d observations $Ysim =(\frac{Xsim_{1}}{Xsim_{2}})$. Don't print the outcome of Ysim (you can only print the first 6 value). (10 points)

You can see below rather than run the entire symulation, we can sample the universe of Y with replacement 10000 times and get the same result.  Then we use the head function with parameter n = 6 on Ysim to display the fist 6 values of Ysim.  I also ran a length function on Ysim to confirm it's size.  We set replace equal to TRUE because you cannon draw 10000 samples from a univers of < 10000 without replacement.

```{r}
# Hint: you can use sample() function in R to simulate 10000 instances of random dice numbers. Do this in R: Xsim1 <- sample(1:6, size = 10000, replace = TRUE), then the same for Xsim2. Now you have 10000 instances of rolling a die twice. Why replace = TRUE? 

Ysim = sample(Y,10000,replace = TRUE)
head(Ysim,n = 6)
length(Ysim)


# Then find Ysim=Xsim1/Xsim2 in R. Now you have 10000 instances of simulated Y values, saved in Ysim. Ysim is a new independently and identically distributed (iid) random variable. Ysim values are independent of each other because sampling is with replacement, or the outcome of the first roll does not have any impact on the outcome of the second roll.

```

c. Draw the graph of successive average (cumulative mean) of $Ysim$. Discuss your observations based on what you see in the graph of successive average and the answer of part a for the mean($Y$). (12 points)

From the plot below we can see that the cumulative mean oscillates greatly around the expected value early in the sumulation, but as the symulation caries on the law of large numbers takes over and the cumulative mean converges to the expected value of 1.429.


```{r}
# For plotting the cumulative mean of Ysim values we use plot(cumMean, type = "l"), but before that, you need to make a vector of cumulative means (cumMean) for Ysim values. Please watch at 2:50 in my first video in Module 2.

sumYsim = cumsum(Ysim)

n = length(sumYsim)

cumMean = c()

for (i in 1:n) {
  cumMean[i] <- sumYsim[i]/i
}

plot(cumMean, type = "l", ylab = "Cumulative Mean", xlab = 'Roll Number',main = "Cumulative Mean of Dice Simulation")



```

# Problem 2 (52 points)
A small hotel has 10 rooms. From experience they know that 20% of the time, people who make reservations do not show up, so as a result, they overbook by accepting 12 reservations for a given night. Let *X* be the number of no shows that night (people who don't show up). 

a. What is the expected number of "no shows" that night? What is the standard deviation of that number? (8 points)

Given 12 bookings and a no show probability of .2, we can expect that 2.4 guests will not show up for their reservation.  The standard deviation is 1.386 no shows.



```{r}
# Hint: X (the number of no shows) is a binomial random variable: X~binomial(p=0.2, n = 12). There is a given formula for finding the expected value and standard deviation of the binomial random variable. Watch Module 4, lecture part 3, when I talk about the properties of binomial random variables. 
n = 12
p = .2

exp_val =n*p
exp_val
vari = n*p*(1-p)
vari
stdev = round(sqrt(vari),3)
stdev


```
b. Is $X$ continuous or discrete. (2 points)

The variable X is discrete.  If we have 12 bookings, we cannot have 11.5 no shows.  We know that X is discrete because there is a finite set of outcomes X ={0,1,2,...12}.

c. What is the range of $X$? (4 points)

The range of X is 12.

d. Find the probability of all possible value of $X$. Round by 3 decimal places. (10 points)

One can see below all the probabilities of X.  The highest probability is at 2 no shows with corresponds with our expected value of 2.4.  We can confirm the validity of our results by summing the probability vector which must sum to 1.  

```{r}
# You need to find P(X=x). You can answer it by using dbinom(X, size = 12, prob=.2) in R.
# Watch Module 4, lecture part 3
X <- 0:12
n <- 12
p <- .2

p_x <- round(dbinom(X,n,p),3)
p_x
sum(p_x)


```

e. Creat a needle plot for all peobability values of part d. (10 points)

You can see from our needle plot that our greates probability lies at 2 no shows which is the descrete value closest to our expected value.

```{r}
# Watch Module 4, lecture part 3
X <- 0:12
n <- 12
p <- .2
p_x <- round(dbinom(X,n,p),3)
plot(X, p_x, type = "h",ylab = 'p(x)', main = 'Needle Plot of X and P(x)')
points(X, p_x, pch = 18)
```

f. What is the probability that the hotel will end up with more customers than they can handle (that is, more people with reservations than available rooms will arrive)? (4 points)

The probability that the hotel will end up with more customers than they can handle is the sum of the probability that there are 0 cancelations and the probability that ther is only 1 cancelation or 27.5 percent.  

```{r}
# Hint: X is the number of no-shows and we know that the p=0.2, and n=12. If zero or one customer does not show up, then the hotel ends up with more customer than available rooms. 
X <- 0:12
n <- 12
p <- .2
p_x <- round(dbinom(X,n,p),3)

p_overbook <- p_x[1]+p_x[2]
p_overbook

```


g. Simulate 10000 instances for the number of no-shows and call them $Xsim$; then estimate the expected value and standard deviation of the number of the simulated no-shows. What explains why the simulation results are different from the answers to a? (14 points)

The simulation results are slightly different than our previous answeres because we are using a sample.  We have generated 10,000 outcomes, the expected value is based on the long run relative frequency of an outcome. We can see in our results, however, that our 10,000 simulations is getting us within hundredths of the expected results. 

```{r}
# Hint: You can generate 10000 (or more) binomial instances based on n=12, and p=0.2 as follows:
# Xsim <-  rbinom(10000, 12, 0.2)
# Now Xsim is a data vector, and we can find its means and stdev by mean() and sd() functions in R.

Xsim <- rbinom(10000, 12, 0.2)
mean(Xsim)
sd(Xsim)

```

```{r}
# Hint: X is the number of no-shows and we know that the p=0.2, and n=12. If zero or one customer does not show up, then the hotel ends up with more customer than available rooms. 
X <- 0:10
n <- 10
p <- .07
p_x <- round(dbinom(X,n,p),4)

p_overbook <- 1-(p_x[1]+p_x[2])
p_overbook

```
